---
title: "WHOI-Plankton Data Integration"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WHOI-Plankton Data Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The [WHOI-Plankton](https://hdl.handle.net/1912/7341) dataset (Sosik et al. 2015) contains millions of microscopic marine plankton images captured by the IFCB, and manually classified into > 100 categories by researchers at the Woods Hole Oceanographic Institution (WHOI). The dataset is part of a larger collection of over 700 million images gathered since 2006 at the Martha's Vineyard Coastal Observatory (MVCO), with real-time and archived data accessible via the [IFCB Data Dashboard](https://ifcb-data.whoi.edu/mvco). Annotated images are stored in annual zip files, organized into class-specific subdirectories. Example images for each class are available on the WHOI-Plankton [GitHub page](https://whoigit.github.io/whoi-plankton/).

`iRfcb` provides functions to interact with the WHOI-Plankton data and the IFCB Dashboard, making it easier to integrate annotated images into custom training datasets. It supports incorporation of these images using MATLAB code from the [`ifcb-analysis`](https://github.com/hsosik/ifcb-analysis) repository (Sosik and Olson 2007), as demonstrated in this tutorial.

## Getting Started

### Installation

You can install the package from CRAN using:
```{r, eval=FALSE}
install.packages("iRfcb")
```

Load the `iRfcb` library:
```{r, eval=FALSE}
library(iRfcb)
```

### Download Sample Data

To get started, download sample data from the [SMHI IFCB Plankton Image Reference Library](https://doi.org/10.17044/scilifelab.25883455.v3) (Torstensson et al. 2024). This dataset will serve as the primary dataset in this example, which we will expand by incorporating training images from the WHOI-Plankton dataset.

```{r, eval=FALSE}
# Define data directory
data_dir <- "data"

# Download and extract test data in the data folder
ifcb_download_test_data(
  dest_dir = data_dir,
  max_retries = 10,
  sleep_time = 30,
  verbose = FALSE
)
```

## Download WHOI-Plankton Data

This section demonstrates how to download the WHOI-Plankton dataset `iRfcb`.

### Define Download Paths

First we define the download directories:

```{r, eval=FALSE}
# Define paths to download destinations
png_folder_whoi <- file.path(data_dir, "whoi_plankton", "png_images", "extracted_images")
raw_folder_whoi <- file.path(data_dir, "whoi_plankton", "data")
manual_folder_whoi <- file.path(data_dir, "whoi_plankton", "manual")
class2use_file_whoi <- file.path(data_dir, "whoi_plankton", "config", "class2use_whoi.mat")
blobs_folder_whoi <- file.path(data_dir, "whoi_plankton", "blobs")
```

### Download and Prepare WHOI-Plankton Data

The [`ifcb_prepare_whoi_plankton`](../reference/ifcb_prepare_whoi_plankton.html) function enables downloading zipped `.png` images from selected years of the [WHOI-Plankton dataset](https://hdl.handle.net/1912/7341), with an option to extract them using the `extract_images` argument. It also retrieves raw data (`.roi`, `.hdr` and `.adc` files), along with blobs and features (when available), from the [MVCO IFCB Dashboard](https://ifcb-data.whoi.edu/mvco). Interactions with the IFCB Dashboard is handled through the [`ifcb_download_dashboard_data`](../reference/ifcb_download_dashboard_data.html) function. Finally, the function generates manual `.mat` files for storing class information, ensuring compatibility with the MATLAB code in the [ifcb-analysis](https://github.com/hsosik/ifcb-analysis) repository (Sosik and Olson 2007).

In this example, we use data from the years 2013 and 2014.

```{r, eval=FALSE}
# Download and prepare the WHOI-Plankton dataset
ifcb_prepare_whoi_plankton(
  years = 2013:2014,
  png_folder = png_folder_whoi, 
  raw_folder = raw_folder_whoi, 
  manual_folder = manual_folder_whoi, 
  class2use_file = class2use_file_whoi, 
  extract_images = FALSE,
  skip_classes = NULL,
  download_blobs = TRUE, # Optionally download blobs
  blobs_folder = blobs_folder_whoi,
  download_features = FALSE,
  quiet = TRUE
)
```

To exclude certain images from the training dataset, either exclude the class completely with the `skip_classes` argument, or set `extract_images = TRUE` and manually delete specific `.png` files from the `png_folder` and rerun `ifcb_prepare_whoi_plankton`.

## Merge Data with Existing Dataset

The [`ifcb_merge_manual`](../reference/ifcb_merge_manual.html) function allows you to integrate data from the WHOI-Plankton dataset with your existing training dataset created using the [ifcb-analysis](https://github.com/hsosik/ifcb-analysis) code (Sosik and Olson 2007). The merged dataset and class2use file will be saved in the locations specified by `manual_folder_output` and `class2use_file_output`.

```{r, eval=FALSE}
# Define paths to existing manual dataset
class2use_file_smhi <- file.path(data_dir, "config", "class2use.mat")
manual_folder_smhi <- file.path(data_dir, "manual")

# Define paths to final merged dataset
class2use_file_merged <- file.path(data_dir, "merged_data", "config", "class2use.mat")
manual_folder_merged <- file.path(data_dir, "merged_data", "manual")

# Merge WHOI-Plankton with existing manual dataset
ifcb_merge_manual(
  class2use_file_base = class2use_file_smhi,
  class2use_file_additions = class2use_file_whoi,
  class2use_file_output = class2use_file_merged,
  manual_folder_base = manual_folder_smhi,
  manual_folder_additions = manual_folder_whoi,
  manual_folder_output = manual_folder_merged,
  quiet = TRUE
)
```

The newly merged dataset is now ready for use in training and testing machine learning models. To integrate it with the [`ifcb-analysis`](https://github.com/hsosik/ifcb-analysis) MATLAB code, blobs and features are extracted for the merged dataset (if not already downloaded through the IFCB Dashboard). Random forest models can then be built using the `ifcb-analysis` code.

This concludes this tutorial for the `iRfcb` package. For more detailed information, refer to the package documentation or the other [tutorials](../articles/index.html). See how data pipelines can be constructed using `iRfcb` in the following [Example Project](https://github.com/nodc-sweden/ifcb-data-pipeline). Happy analyzing!

## Citation

```{r, echo=FALSE}
# Print citation
citation("iRfcb")
```

## References
- Sosik, H. M. and Olson, R. J. (2007) Automated taxonomic classification of phytoplankton sampled with imaging-in-flow cytometry. Limnol. Oceanogr: Methods 5, 204–216.
- Sosik, H. M., Peacock, E. E. and Brownlee E. F. (2015), Annotated Plankton Images - Data Set for Developing and Evaluating Classification Methods. https://doi.org/10.1575/1912/7341
- Torstensson, A., Skjevik, A-T., Mohlin, M., Karlberg, M. and Karlson, B. (2024). SMHI IFCB Plankton Image Reference Library. SciLifeLab. Dataset. https://doi.org/10.17044/scilifelab.25883455.v3
