---
title: "Metadata template"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Metadata template}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{LivingNorwayR}
---

```{r, include = FALSE}
if (!requireNamespace("LivingNorwayR", quietly = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
  message("LivingNorwayR is not available, so this vignette will not run.")
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, include=FALSE}
# This is setup code that is not displayed in the rendered HTML file but includes packages
# needed for producing the metadata document
knitr::opts_chunk$set(echo = TRUE)
# Check if LivingNorwayR is available
LN_installed <- requireNamespace("LivingNorwayR", quietly = TRUE)

# Load LivingNorwayR conditionally
if (LN_installed) {
  library(LivingNorwayR)
} else {
  message("LivingNorwayR is not available, skipping related code.")
}
```

This R Markdown document is based on a template provided by the [`LivingNorwayR`](https://livingnorway.github.io/LivingNorwayR/) package, and is used for creating a `.eml` file in `vignette("dwca-tutorial")`.

When documenting project metadata in Markdown, the `LivingNorwayR` package provides various functions to flag specific text sections for export into an EML file. The code chunk below highlights these flagged sections in red when rendered in HTML output. However, for standard metadata descriptions, this highlighting is typically unnecessary, so you should remove this code chunk when creating your own metadata descriptions.

```{css}
span.LNmetadata {
  color: red;
}
```

## The Dataset

Metadata must include a dataset tag, which requires an ID. This ID acts as a parent identifier for numerous other tags that describe the dataset. In many cases, you may not want any text associated with this tag to appear in the output. To achieve this, set the `isHidden` argument to `TRUE`, making the tag invisible in the rendered output. The primary purpose of this function is to assign an ID to the `dataset` tag (e.g., **iRfcbDataset**), enabling other elements to reference it. Additionally, you can provide a title for the dataset using the `title.tagText` argument within the function.

`r if (LN_installed) {LNdataset("", tagID = "iRfcbDataset", isHidden = TRUE, title.tagText = "iRfcb Test Data")}`

We can associate keywords with the dataset by creating a `keywordSet` tag using the appropriate tagging function. For example: `r if (LN_installed) {LNkeywordSet("", tagID = "iRfcbKeywordSet", parentID = "iRfcbDataset", isHidden = TRUE)}`. Next, individual keywords can be specified within this keywordSet using the LNkeyword function. Examples include: `r if (LN_installed) {LNkeyword("Imaging FlowCytobot", parentID = "iRfcbKeywordSet")}`, `r if (LN_installed) {LNkeyword("phytoplankton", parentID = "iRfcbKeywordSet")}`, `r if (LN_installed) {LNkeyword("microzooplankton", parentID = "iRfcbKeywordSet")}`, `r if (LN_installed) {LNkeyword("imaging", parentID = "iRfcbKeywordSet")}` and `r if (LN_installed) {LNkeyword("sampling event", parentID = "iRfcbKeywordSet")}`.

We must also specify some contact information for the individual or organization responsible for coordinating with users of the dataset.  Here the responsible user is `r if (LN_installed) {LNcontact("", tagID = "iRfcbContact", parentID = "iRfcbDataset")}` `r if (LN_installed) {LNindividualName("", parentID = "iRfcbContact", givenName.tagText = "Anders", surName.tagText = "Torstensson")}`. This approach ensures that all keywords are grouped under the `iRfcbKeywordSet` tag, which is linked to the dataset through its `parentID`.

## Abstract

An abstract for the dataset is required. You can flag the abstract for export to EML using the following inline code:

`r if (LN_installed) {LNabstract("Imaging FlowCytobot sample data from the iRfcb R package, collected onboard R/V Svea between 2022 and 2023.", tagID = "iRfcbAbstract", parentID = "iRfcbDataset")}`

The `LivingNorwayR` package also supports adding alternative translations for EML elements. In the inline code above, we specified the `tagID` argument. To provide an alternative translation for the element associated with that `tagID`, you can use the following code:

`r if (LN_installed) {LNaddTranslation("Exempeldata från Imaging FlowCytobot i iRfcb R-paketet, insamlade ombord R/V Svea mellan 2022 och 2023.", lang = "nb", parentID = "iRfcbAbstract")}`

By default, alternative translations are hidden in the rendered HTML output. While the information is included in the file, it is not visible when viewed in a browser. This is useful when you want the translation to be exported to the EML file without appearing in the rendered HTML. If you prefer the alternative translation to be displayed, simply add the argument `isHidden = FALSE` to the `LNaddTranslation` function.

## Dataset creators

The dataset was created by the following people:

  + `r if (LN_installed) {LNcreator(tagText = "", tagID = "iRfcbCreator", parentID = "iRfcbDataset", individualName.tagText = "", individualName.givenName.tagText = "Anders", individualName.surName.tagText = "Torstensson")}` who is a `r if (LN_installed) {LNpositionName(parentID = "iRfcbCreator", tagText = "researcher")}` at the `r if (LN_installed) {LNorganizationName(parentID = "iRfcbCreator", tagText = "Swedish Meteorological and Hydrological Institute")}` (`r if (LN_installed) {LNelectronicMailAddress(parentID = "iRfcbCreator", tagText = "anders.torstensson@smhi.se")}`).

