---
title: "Quality Control of IFCB Data"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality Control of IFCB Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette provides an overview of quality control (QC) methods for Imaging FlowCytobot (IFCB) data using the `iRfcb` package. The package offers tools to analyze Particle Size Distribution (PSD) following Hayashi et al. (2025), verify geographical positions, and integrate contextual data from sources like ferrybox systems. These QC workflows ensure high-quality datasets for phytoplankton and microzooplankton monitoring in marine ecosystems.

You'll learn how to:

- Set up the `iRfcb` package and `Python` environment.
- Run native-R integrity checks on raw sample files (the `.hdr`/`.adc`/`.roi` triplet).
- Analyze particle size distributions for data quality.
- Check spatial metadata for proximity to land, basin classification, and missing positions.

Follow this tutorial to streamline the QC process and ensure reliable IFCB data.

## Getting Started

### Installation

You can install the package from CRAN using:
```{r, eval=FALSE}
install.packages("iRfcb")
```
Some functions from the `iRfcb` package used in this tutorial require `Python` to be installed. You can download `Python` from the official website: [python.org/downloads](https://www.python.org/downloads/).

The `iRfcb` package can be configured to automatically activate an installed Python virtual environment (venv) upon loading by setting an environment variable. For more details, please refer to the package [README](https://europeanifcbgroup.github.io/iRfcb/).

Load the `iRfcb` library:
```{r}
library(iRfcb)
```

```{r, include=FALSE}
library(reticulate)

# Define path to virtual environment
env_path <- file.path(tempdir(), "iRfcb") # Or your preferred venv path

# Install python virtual environment
tryCatch({
  ifcb_py_install(envname = env_path)
}, error = function(e) {
  warning("Python environment could not be installed.")
})
```

```{r, echo=FALSE}
# Check if Python is available
if (!py_available(initialize = TRUE)) {
  knitr::opts_chunk$set(eval = FALSE)
  warning("Python is not available. Skipping vignette evaluation.")
} else {
  # List available packages
  available_packages <- py_list_packages(python = reticulate::py_discover_config()$python)
  
  # Check if pandas and matplotlib are available
  if (!"pandas" %in% available_packages$package ||
      !"matplotlib" %in% available_packages$package) {
    knitr::opts_chunk$set(eval = FALSE)
    warning("Required python modules are not available. Skipping vignette evaluation.")
  }
}
```

### Download Sample Data

To get started, download sample data from the [SMHI IFCB Plankton Image Reference Library](https://doi.org/10.17044/scilifelab.25883455.v3) (Torstensson et al. 2024) with the following function:
```{r, eval=FALSE}
# Define data directory
data_dir <- "data"

# Download and extract test data in the data folder
ifcb_download_test_data(
  dest_dir = data_dir,
  verbose = FALSE
)
```

```{r, include=FALSE}
# Define data directory
data_dir <- "data"

# Download and extract test data in the data folder
if (!dir.exists(data_dir)) {
  # Download and extract test data if the folder does not exist
  ifcb_download_test_data(
    dest_dir = data_dir,
    verbose = FALSE
  )
}
```

## Sample Integrity Checks

Before any feature-based analysis, the raw sample files themselves can be checked for completeness and consistency. `ifcb_qc_sample()` validates the IFCB file triplet (`.hdr`, `.adc`, and `.roi`) directly and returns a tidy tibble with one row per sample. Unlike the particle size distribution check below, it runs in native R and requires **no Python** and **no pre-computed features**, which makes it a quick first-pass screen for missing files, truncated or aborted runs, and metadata anomalies.

```{r}
# Run sample-level QC on the raw data folder
qc <- ifcb_qc_sample("data/data/2023")

# Inspect the integrity flags
qc[, c("sample", "n_rois", "hdr_roi_count", "ml_analyzed", "syringe_ml", "qc_pass")]
```

The overall `qc_pass` column summarizes five integrity checks: the triplet is complete, the number of imaged ROIs matches the header `roiCount`, the `.roi` file is not truncated, the header and ADC agree on the run time, and the analyzed volume is sane (by default bounded by the header `SyringeSampleVolume`, reported as `syringe_ml`). To keep only the samples that pass:

```{r, eval=FALSE}
# Keep clean, non-bead samples for downstream analysis
dplyr::filter(qc, qc_pass, !is_bead_run)
```

Several optional, advisory flags can be enabled by supplying thresholds. These do not affect `qc_pass`, but let you screen for overloaded or environmentally compromised runs:

```{r, eval=FALSE}
# Flag oversized .roi files (MB), high humidity (%), and high temperature
ifcb_qc_sample(
  "data/data/2023",
  max_roi_mb = 5,
  max_humidity = 70,
  max_temperature = 35
)
```

For a deeper, statistical quality check based on the particle size distribution, continue with the section below.

## Particle Size Distribution

IFCB data can be quality controlled by analyzing the particle size distribution (PSD) (Hayashi et al. 2025). `iRfcb` uses the code available at [https://github.com/kudelalab/PSD](https://github.com/kudelalab/PSD), which is efficient in detecting samples with bubbles, beads, incomplete runs etc. The PSD analysis operates on the feature files (`<bin>_features_v4.csv`) supplied via `feature_folder`. If you have not yet generated these features, they can be computed from raw data in R with `ifcb_extract_features()` (see `vignette("introduction")`). Before running the PSD quality check, ensure the necessary Python environment is set up and activated:

```{r, eval=FALSE}
# Define path to virtual environment
env_path <- "~/.virtualenvs/iRfcb" # Or your preferred venv path

# Install python virtual environment
ifcb_py_install(envname = env_path)

# Run PSD quality control
psd <- ifcb_psd(
  feature_folder = "data/features/2023",
  hdr_folder = "data/data/2023",
  save_data = FALSE,
  output_file = NULL,
  plot_folder = NULL,
  use_marker = FALSE,
  start_fit = 10,
  r_sqr = 0.5,
  beads = 10 ** 12,
  bubbles = 150,
  incomplete = c(1500, 3),
  missing_cells = 0.7,
  biomass = 1000,
  bloom = 5,
  humidity = 70
)
```

```{r, include=FALSE}
# Run PSD quality control
psd <- ifcb_psd(
  feature_folder = "data/features/2023",
  hdr_folder = "data/data/2023",
  save_data = FALSE,
  output_file = NULL,
  plot_folder = NULL,
  use_marker = FALSE,
  start_fit = 10,
  r_sqr = 0.5,
  beads = 10 ** 12,
  bubbles = 150,
  incomplete = c(1500, 3),
  missing_cells = 0.7,
  biomass = 1000,
  bloom = 5,
  humidity = 70,
  fea_v = 2, # Use v2 features
)
```

The results can be printed and visualized through plots:
```{r}
# Print output from PSD
head(psd$fits)
head(psd$flags)

# Plot PSD of the first sample
plot <- ifcb_psd_plot(
  sample_name = psd$data$sample[1],
  data = psd$data,
  fits = psd$fits,
  start_fit = 10
)

# Print the plot
print(plot)
```

## Geographical QC/QA

### Check If IFCB Is Near Land

To determine if the IFCB is near land (i.e. ship in harbor), examine the position data in the `.hdr` files (or from vectors of latitudes and longitudes):

```{r}
# Read HDR data and extract GPS position (when available)
gps_data <- ifcb_read_hdr_data(
  "data/data/", 
  gps_only = TRUE, 
  verbose = FALSE # Do not print progress bar
)

# Create new column with the results
gps_data$near_land <- ifcb_is_near_land(
  gps_data$gpsLatitude,
  gps_data$gpsLongitude,
  distance = 100, # 100 meters from shore
  shape = NULL # Using the default NE 1:10m Land Polygon
) 

# Print output
head(gps_data)

# Alternatively, you can choose to plot the points on a map
near_land_plot <- ifcb_is_near_land(
  gps_data$gpsLatitude,
  gps_data$gpsLongitude,
  distance = 2500, # 2500 meters from shore
  plot = TRUE,
)

# Print the plot
print(near_land_plot)
```

For more accurate determination, a detailed coastline `.shp` file may be required (e.g. the [EEA Coastline Polygon](https://www.eea.europa.eu/data-and-maps/data/eea-coastline-for-analysis-2/gis-data/eea-coastline-polygon)). Refer to the help pages of `ifcb_is_near_land()` for further information.

### Check Which Sub-Basin an IFCB Sample Is From

To identify the specific sub-basin of the Baltic Sea (or using a custom shape-file) from which an IFCB sample was collected, analyze the position data:
```{r}
# Define example latitude and longitude vectors
latitudes <- c(55.337, 54.729, 56.311, 57.975)
longitudes <- c(12.674, 14.643, 12.237, 10.637)

# Check in which Baltic sea basin the points are in
points_in_the_baltic <- ifcb_which_basin(latitudes, longitudes, shape_file = NULL)
# Print output
print(points_in_the_baltic)

# Plot the points and the basins
ifcb_which_basin(latitudes, longitudes, plot = TRUE, shape_file = NULL)
```

This function reads a pre-packaged shapefile of the Baltic Sea, Kattegat, and Skagerrak basins from the `iRfcb` package by default, or a user-supplied shapefile if provided. The shapefiles provided in `iRfcb` originate from [SHARK](https://shark.smhi.se/en/). 

### Check If Positions Are Within the Baltic Sea or Elsewhere

This check is useful if you only want to apply a classifier specifically to phytoplankton from the Baltic Sea.

```{r}
# Define example latitude and longitude vectors
latitudes <- c(55.337, 54.729, 56.311, 57.975)
longitudes <- c(12.674, 14.643, 12.237, 10.637)

# Check if the points are in the Baltic Sea Basin
points_in_the_baltic <- ifcb_is_in_basin(latitudes, longitudes)

# Print results
print(points_in_the_baltic)

# Plot the points and the basin
ifcb_is_in_basin(latitudes, longitudes, plot = TRUE)
```

This function reads a land-buffered shapefile of the Baltic Sea Basin from the `iRfcb` package by default, or a user-supplied shapefile if provided.

### Find Missing Positions from RV Svea Ferrybox

This function is used by SMHI to collect and match stored ferrybox positions when they are not available in the `.hdr` files. An example ferrybox data file is provided in `iRfcb` with data matching sample **D20220522T000439_IFCB134**.

```{r}
# Print available coordinates from .hdr files
head(gps_data, 4)

# Define path where ferrybox data are located
ferrybox_folder <- "data/ferrybox_data"

# Get GPS position from ferrybox data
positions <- ifcb_get_ferrybox_data(gps_data$timestamp, 
                                    ferrybox_folder)

# Print result
head(positions)
```

### Find Contextual Ferrybox Data from RV Svea

The `ifcb_get_ferrybox_data()` function can also be used to extract additional ferrybox parameters, such as temperature (parameter number **8180**) and salinity (parameter number **8181**). 

```{r}
# Get salinity and temperature from ferrybox data
ferrybox_data <- ifcb_get_ferrybox_data(gps_data$timestamp,
                                        ferrybox_folder,
                                        parameters = c("8180", "8181"))

# Print result
head(ferrybox_data)
```

This concludes this tutorial for the `iRfcb` package. For additional guides—such as data sharing and MATLAB integration—please refer to the other tutorials available on the project's [webpage](https://europeanifcbgroup.github.io/iRfcb/). See how data pipelines can be constructed using `iRfcb` in the following [Example Project](https://github.com/nodc-sweden/ifcb-data-pipeline). Happy analyzing!

```{r, include=FALSE}
# Clean up
unlink(data_dir, recursive = TRUE)
unlink(env_path, recursive = TRUE)
```

## Citation

```{r, echo=FALSE}
# Print citation
citation("iRfcb")
```

## References
- Hayashi, K., Enslein, J., Lie, A., Smith, J., Kudela, R.M., 2025. Using particle size distribution (PSD) to automate imaging flow cytobot (IFCB) data quality in coastal California, USA. International Society for the Study of Harmful Algae. https://doi.org/10.15027/0002041270
- Torstensson, A., Skjevik, A-T., Mohlin, M., Karlberg, M. and Karlson, B. (2024). SMHI IFCB Plankton Image Reference Library. SciLifeLab. Dataset. https://doi.org/10.17044/scilifelab.25883455.v3
